---
title: "CodeBook"
author: "ISY"
date: "12/13/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Assignment for Getting and Cleaning Data Course- Week 4**

In this markdown document, I will describe the steps that I took while working
on "run_analysis.R" file, which you can find under my Github repository.

*Step 1: Acquiring the zip file*

In the first step, we will be downloading the data that we will use in the R script.

```{r}
#Define the url first
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"

#Define destination
destination <- "/Users/ihsanselcukyurttas/Documents/TestingGIT/getcleandata/data.zip"

#download file
download.file(url, destfile = destination, method = "curl")

#Unzip file

unzip(destination)

```

*Step 2: Loading text files to RStudio Environment*

The assignment asks us to create a data table that merges all the information
that comes from the mobile phone. As we look at the folder that we have created
a moment ago by unziping the zip file that contains the data, we can see that
values are contained in ".txt" files. We will call each of them and assign them
to new variables.

```{r}
# Defining directories to variables

#Defining general directories
directory0 <- "/Users/ihsanselcukyurttas/Documents/TestingGIT/getcleandata/UCI HAR Dataset"
training <- file.path(directory0, "train")
test <- file.append(directory, "test")

# Creating training variables

trainSubjects <- read.table(file=file.path(training,"subject_train.txt"))
trainValues <- read.table(file=file.path(training, "X_train.txt"))
trainActivity <- read.table(file= file.path(training, "y_train.txt"))

#Creating testing variables
testSubjects <- read.table(file=file.path(test, "subject_test.txt"))
testValues <- read.table(file=file.path(test, "X_test.txt"))
testActivity <- read.table(file.path(test, "y_test.txt"))

#Creating features variable
features <- read.table(file = file.path(directory0, "features.txt"))

#Creating activity label variable
activity <- read.table(file= file.path(directory0, "activity_labels.txt"))

```

*Step 3: Creating Intermediary Data Frames*

In this step, we will be creating a merged data frame with the variables that
we have defined a moment ago. Recall from the readme.txt document,
each variable has its unique name that is stored in "features.txt" document.
We will assign those variable names to our merged data table. 

```{r}
#Load data.table package to create data tables
library(data.table)

#Combine all

humanActivity <- rbind(
     cbind(trainSubjects, trainActivity, trainValues),
     cbind(testSubjects,testActivity, testValues)
)

#Rename column names

colnames(humanActivity) <- c("Subject", "Activity", features[,2])

```

*Step 4: Selecting expected columns*

In this step,  we will select the variables that shows us means and standard 
deviations. We will use regular expressions to manipulate strings of
column names. For a quick review, you can check regular expressions cheat sheet,
which can be downloaded from RStudio's respective website. I am also leaving a
download link below.

Link: https://github.com/rstudio/cheatsheets/raw/master/regex.pdf

```{r}
#Create a logical vector to assess which columns will be selected
colExtract <- grepl(("Subject|Activity|mean|std"), colnames(humanActivity))

#Apply the logical vector to combined values dataset to extract the
#expected columns

humanActivity <- humanActivity[,colExtract]

```

*Step 5: Changing the activites with respective string type values*

As "activities.txt" document tells us, each number in "Activity" column of 
"humanActivity" dataset means a factor. Let's change those numbers to their respective
factors.

```{r}
humanActivity$Activity <- factor(humanActivity$Activity, 
                                 levels = activity[, 1], labels = activity[, 2])
```


*Step 6: Change variable names to descriptive variable names*

Finally, our only job right now is to change variable names to more descriptive.
Remember that features dataset's second column didn't have that much of an aesthetic
look. Lets change them. We will use regular expressions and gsub() function here.
To have a quick review, I recommend checking regular expressions cheatsheet which
I have added its website link above.

```{r}
names <- colnames(humanActivity)

#Remove the special characters
names <- gsub("[\\(\\)-]", "", names)

#Rename all

names <- gsub("^f", "frequencyDomain", names)
names <- gsub("^t","timeDomain", names)
names <- gsub("Acc","Accelerometer", names)
names <- gsub("Gyro", "Gyroscope", names)
names <- gsub("Freq", "Frequency", names)
names <- gsub("mean", "Mean", names)
names <- gsub("std", "SD", names)

#Use new names list for humanActivity column names
colnames(humanActivity) <- names
```

*Step 7: Create and extract a dataset with the average of each variable*
*for each subject and each activity*

```{r}
#Create a dataset first
tidy_dataset <- humanActivity %>% 
     group_by(Subject, Activity) %>%
     summarize_each(funs(mean))
```

You may ask about the last line. Get the help page of summarize_each() and funs().

```{r}
write.table(tidy_dataset, file = "/Users/ihsanselcukyurttas/Documents/TestingGIT/getcleandata",
            row.names = FALSE)
```



